Denotational Semantics for SSA with Weak Memory Operations

### 2023-10-18 pldi24/paper.tex
Tweaked substitution/semantic substitution, text stubs

% Syntax
\newcommand{\letexpr}[3]{\ensuremath{\ms{let}\;#1 = #2\;\ms{in}\;#3}}
\newcommand{\letstmt}[3]{\ensuremath{\ms{let}\;#1 = #2; #3}}
\newcommand{\brb}[2]{\ms{br}\;#1\;#2}
\newcommand{\lbrb}[2]{\brb{\lbl{#1}}{#2}}
\newcommand{\ite}[3]{\ms{if}\;#1\;\{#2\}\;\ms{else}\;\{#3\}}
\newcommand{\ewhere}[2]{\ms{then}\;#1\;\ms{where}\;#2}
\newcommand{\where}[2]{#1\;\ms{where}\;#2}
\newcommand{\wbranch}[3]{#1(#2) \lto #3}
\newcommand{\lwbranch}[3]{\wbranch{\lbl{#1}}{#2}{#3}}
\newcommand{\bsplice}[3]{#1(#2)\;\{#3\}}
\newcommand{\lbsplice}[3]{\bsplice{\lbl{#1}}{#2}{#3}}
\newcommand{\csplits}[3]{#1 \mapsto #2;#3}
\newcommand{\cwk}[2]{#1 \mapsto #2}
\newcommand{\lwk}[2]{#1 \rightsquigarrow #2}
\newcommand{\tlin}[2]{\ms{lin}_{#2}(#1)}
\newcommand{\ltlin}[3]{\ms{lin}_{#3}(#1^{#2})}
\newcommand{\thyp}[3]{#1: {#2}^{#3}}
\newcommand{\lhyp}[3]{#1[#2](#3)}
\newcommand{\llhyp}[3]{\lhyp{\lbl{#1}}{#2}{#3}}
\newcommand{\rle}[1]{{\scriptsize\textsf{#1}}}
\newcommand{\taff}{\ms{a}}
\newcommand{\trel}{\ms{r}}
\newcommand{\tint}{\infty}
\newcommand{\hasty}[5]{#1 \vdash_{#2} #3: {#4}^{#5}}
\newcommand{\haslb}[3]{#1 \vdash #2 \rhd #3}
\newcommand{\lhaslb}[3]{#1 \vdash #2 \rhd #3}
\newcommand{\issubst}[3]{#1: #2 \mapsto #3}
\newcommand{\lbsubst}[3]{#1: #2 \rightsquigarrow #3}
\newcommand{\exprletsubst}[2]{{#1};{#2}}
\newcommand{\stmtletsubst}[2]{{#1};{#2}}
\newcommand{\mhole}[1]{{#1}^?}
\newcommand{\lhole}[1]{?#1}
\newcommand{\mhasty}[6]{#1;#2 \vdash_{#3} #4: {#5}^{#6}}
\newcommand{\mhaslb}[4]{#1;#2 \vdash #3 \rhd #4}
\newcommand{\mlhaslb}[4]{#1;#2 \vdash #3 \rhd #4}
\newcommand{\tyhole}[5]{#1: #2 \mapsto_{#3} {#4}^{#5}}
\newcommand{\blkhole}[3]{#1: #2 \mapsto #3}
\newcommand{\cfghole}[3]{#1: #2 \mapsto #3}
\newcommand{\substctx}[2]{{#1}^{#2}}
\newcommand{\substlbs}[2]{{#1}^{#2}}
\newcommand{\restrictsubst}[2]{{#1}_{#2}}
\newcommand{\subsubst}[2]{#1 \subseteq #2}
\newcommand{\isrw}[3]{#1: #2 \mapsto #3}
\newcommand{\strictlbsubst}[3]{#1: #2 \rightsquigarrow_= #3}

@@

@@

@@

@@

\begin{figure}
  \begin{center}        
    \begingroup
    \renewcommand{\arraystretch}{1.5}
    \setlength{\tabcolsep}{2em}
    \begin{tabular}{rl}
        \multicolumn{1}{c}{Judgment} & \multicolumn{1}{c}{Meaning} \\ \hline
        \(\issubst{\gamma}{\Theta}{\Gamma}\) &
        ... \\
        \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) &
        ... \\
        \(\mhasty{H}{\Gamma}{p}{\mhole{a}}{A}{q}\) &
        ... \\
        \(\mhaslb{H}{\Gamma}{\mhole{t}}{\ms{L}}\) &
        ... \\
        \(\mlhaslb{H}{\ms{L}}{\mhole{L}}{\ms{K}}\) &
        ...
    \end{tabular}
    \endgroup
  \end{center}
  \caption{Metatheoretic typing judgements for \isotopessa}
  \Description{Metatheoretic typing judgements for isotope-SSA}
  \label{fig:ssa-meta-judgements}
\end{figure}

@@

\begin{figure}
  \begin{gather*}
    \boxed{\gamma: \ms{Var} \to \ms{Expr}}
    \\
    \prftree[r]{\rle{subst-nil}}
      {\issubst{\gamma}{\cdot}{\cdot}}
      \qquad
    \prftree[r]{\rle{subst-cons}}
      {\issubst{\gamma}{\Theta_\Gamma}{\Gamma}}
      {\hasty{\Theta_x}{0}{\gamma(x)}{A}{q}}
      {\tlin{\Theta_x}{q}}
      {\csplits{\Theta}{\Theta_\Gamma}{\Theta_x}}
      {\issubst{\gamma}{\Theta}{\Gamma, \thyp{x}{A}{q}}}
    \\
    \boxed{\mc{L}: \ms{Label} \to \ms{Expr} \to \ms{Block} \text{ s.t. for fresh } x, [a/x]\mc{L}(\lbl{\ell}, x) = \mc{L}(\lbl{\ell}, a)}
    \\
    \prftree[r]{\rle{lb-subst-nil}}
    {\lbsubst{\mc{L}}{\cdot}{\ms{K}}}
      \qquad
    \prftree[r]{\rle{lb-subst-cons}}
      {\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
      {\haslb{\Gamma, \thyp{x}{A}{}}{\mc{L}(\lbl{\ell}, x)}{\ms{K}}}
      {\lbsubst{\mc{L}}{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K}}}
    \\
    \prftree[r]{\rle{lb-strict}}
      {\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
      {\ms{labels}(\ms{L}) = \ms{labels}(\ms{K})}
      {\strictlbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
  \end{gather*}
  \caption{Typing rules for \isotopessa substitutions. We assume that for all but finitely many \(x\), \(\lbl{\ell}\), \(\gamma(x) = x\) and \(\mc{L}(\lbl{\ell}, x) = \lbrb{\ell}{x}\).}
  \Description{Typing rules for isotope-SSA substitutions}
  \label{fig:ssa-subst-typing}
\end{figure}

@@

@@

\begin{theorem}[Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\)
    \item If \(\haslb{\Gamma}{t}{\ms{L}}\), \(\strictlbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}\), and then \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\)
  \end{itemize}
\end{theorem}

@@

@@

@@

@@

@@

\begin{theorem}[Rewriting]
  Given \(\isrw{\mc{H}}{H}{I}\), we have
  \begin{itemize}
    \item If \(\mhasty{H}{\Gamma}{p}{a}{A}{q}\), then \(\mhasty{I}{\Gamma}{p}{[\mc{H}]a}{A}{q}\)
    \item If \(\mhaslb{H}{\Gamma}{t}{\ms{L}}\), then \(\mhaslb{I}{\Gamma}{[\mc{H}]t}{\ms{L}}\)
  \end{itemize}
  Furthermore, we have that 
  \begin{itemize}
    \item \(\mhasty{\cdot}{\Gamma}{p}{a}{A}{q} \iff \hasty{\Gamma}{p}{a}{A}{q}\)
    \item \(\mhaslb{\cdot}{\Gamma}{t}{\ms{L}} \iff \haslb{\Gamma}{t}{\ms{L}}\)
  \end{itemize}
\end{theorem}

@@

\TODO{premonoidal categories}

@@

@@

@@

@@

@@

@@

@@

\begin{lemma}[Semantic Weakening] 
  If \(\cwk{\Gamma}{\Delta}\) and \(\hasty{\Gamma}{p}{a}{A}{q}\), then
  \[
    \dnt{\hasty{\Gamma}{p}{a}{A}{q}} = \dnt{\cwk{\Gamma}{\Delta}};\dnt{\hasty{\Delta}{p}{a}{A}{q}}
  \]
\end{lemma}

@@

\begin{lemma}[Substitution Coherence] \
  \begin{itemize}
    \item Given any two derivations \(D_1: \issubst{\gamma}{\Theta}{\Gamma}\), \(D_2: \issubst{\gamma'}{\Theta}{\Gamma}\) such that \(\forall x \in \Theta, \gamma(x) = \gamma'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
    \item Given any two derivations \(D_1: \lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), \(D_2: \lbsubst{\mc{L}'}{\ms{L}}{\ms{K}}\) such that \(\forall \lbl{\ell} \in \ms{L}, \mc{L}(x) = \mc{L}'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
    \item Given any two derivations \(D_1: \strictlbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), \(D_2: \strictlbsubst{\mc{L}'}{\ms{L}}{\ms{K}}\) such that \(\forall \lbl{\ell} \in \ms{L}, \mc{L}(x) = \mc{L}'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
  \end{itemize}
\end{lemma}

@@

\begin{theorem}[Semantic Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item \(\upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{p}
      ;\dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
      = \dnt{\hasty{\Theta}{p}{[\gamma]a}{A}{q}}\)
    \item Given \(\strictlbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}\), \(\upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{}
      ; \dnt{\haslb{\Gamma}{t}{\ms{L}}}
      = \dnt{\haslb{\Theta}{[\gamma]t}{\ms{K}}} 
      ; \dnt{\strictlbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}}\)
  \end{itemize}
  Similarly, given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma}{t}{\ms{L}}\), we have
  \[
    \dnt{\haslb{\Gamma}{t}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
    = \dnt{\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}}
  \]
\end{theorem}

@@

@@

@@

@@

>>>>

% Syntax
\newcommand{\letexpr}[3]{\ensuremath{\ms{let}\;#1 = #2\;\ms{in}\;#3}}
\newcommand{\letstmt}[3]{\ensuremath{\ms{let}\;#1 = #2; #3}}
\newcommand{\brb}[2]{\ms{br}\;#1\;#2}
\newcommand{\lbrb}[2]{\brb{\lbl{#1}}{#2}}
\newcommand{\ite}[3]{\ms{if}\;#1\;\{#2\}\;\ms{else}\;\{#3\}}
\newcommand{\ewhere}[2]{\ms{then}\;#1\;\ms{where}\;#2}
\newcommand{\where}[2]{#1\;\ms{where}\;#2}
\newcommand{\wbranch}[3]{#1(#2) \lto #3}
\newcommand{\lwbranch}[3]{\wbranch{\lbl{#1}}{#2}{#3}}
\newcommand{\bsplice}[3]{#1(#2)\;\{#3\}}
\newcommand{\lbsplice}[3]{\bsplice{\lbl{#1}}{#2}{#3}}
\newcommand{\csplits}[3]{#1 \mapsto #2;#3}
\newcommand{\cwk}[2]{#1 \mapsto #2}
\newcommand{\lwk}[2]{#1 \rightsquigarrow #2}
\newcommand{\tlin}[2]{\ms{lin}_{#2}(#1)}
\newcommand{\ltlin}[3]{\ms{lin}_{#3}(#1^{#2})}
\newcommand{\thyp}[3]{#1: {#2}^{#3}}
\newcommand{\lhyp}[3]{#1[#2](#3)}
\newcommand{\llhyp}[3]{\lhyp{\lbl{#1}}{#2}{#3}}
\newcommand{\rle}[1]{{\scriptsize\textsf{#1}}}
\newcommand{\taff}{\ms{a}}
\newcommand{\trel}{\ms{r}}
\newcommand{\tint}{\infty}
\newcommand{\hasty}[5]{#1 \vdash_{#2} #3: {#4}^{#5}}
\newcommand{\haslb}[3]{#1 \vdash #2 \rhd #3}
\newcommand{\lhaslb}[3]{#1 \vdash #2 \rhd #3}
\newcommand{\issubst}[3]{#1: #2 \mapsto #3}
\newcommand{\lbsubst}[3]{#1: #2 \rightsquigarrow #3}
\newcommand{\exprletsubst}[2]{{#1};{#2}}
\newcommand{\stmtletsubst}[2]{{#1};{#2}}
\newcommand{\mhole}[1]{{#1}^?}
\newcommand{\lhole}[1]{?#1}
\newcommand{\mhasty}[6]{#1;#2 \vdash_{#3} #4: {#5}^{#6}}
\newcommand{\mhaslb}[4]{#1;#2 \vdash #3 \rhd #4}
\newcommand{\mlhaslb}[4]{#1;#2 \vdash #3 \rhd #4}
\newcommand{\tyhole}[5]{#1: #2 \mapsto_{#3} {#4}^{#5}}
\newcommand{\blkhole}[3]{#1: #2 \mapsto #3}
\newcommand{\cfghole}[3]{#1: #2 \mapsto #3}
\newcommand{\substctx}[2]{{#1}^{#2}}
\newcommand{\substlbs}[2]{{#1}^{#2}}
\newcommand{\restrictsubst}[2]{{#1}_{#2}}
\newcommand{\subsubst}[2]{#1 \subseteq #2}
\newcommand{\isrw}[3]{#1: #2 \mapsto #3}
% \newcommand{\strictlbsubst}[3]{#1: #2 \rightsquigarrow_= #3}

@@

\TODO{some text about each typing judgement}

@@

\TODO{some text about each structural rule}

@@

\TODO{some text about each typing rule}

@@

\TODO{some text about each block typing rule}

\subsection{Syntactic Metatheory}

@@

\begin{figure}
  \begin{center}        
    \begingroup
    \renewcommand{\arraystretch}{1.5}
    \setlength{\tabcolsep}{2em}
    \begin{tabular}{rl}
        \multicolumn{1}{c}{Judgment} & \multicolumn{1}{c}{Meaning} \\ \hline
        \(\issubst{\gamma}{\Theta}{\Gamma}\) &
        ... \\
        \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) &
        ... \\
        \(\mhasty{H}{\Gamma}{p}{\mhole{a}}{A}{q}\) &
        ... \\
        \(\mhaslb{H}{\Gamma}{\mhole{t}}{\ms{L}}\) &
        ... \\
        \(\mlhaslb{H}{\ms{L}}{\mhole{L}}{\ms{K}}\) &
        ... \\
        \(\isrw{\mc{H}}{I}{H}\) &
        ... \\
    \end{tabular}
    \endgroup
  \end{center}
  \caption{Metatheoretic typing judgements for \isotopessa}
  \Description{Metatheoretic typing judgements for isotope-SSA}
  \label{fig:ssa-meta-judgements}
\end{figure}

\TODO{some text about each metatheoretic judgement}

\begin{figure}
  \begin{gather*}
    \boxed{\gamma: \ms{Var} \to \ms{Expr}}
    \\
    \prftree[r]{\rle{subst-nil}}
      {\issubst{\gamma}{\cdot}{\cdot}}
      \qquad
    \prftree[r]{\rle{subst-cons}}
      {\issubst{\gamma}{\Theta_\Gamma}{\Gamma}}
      {\hasty{\Theta_x}{0}{\gamma(x)}{A}{q}}
      {\tlin{\Theta_x}{q}}
      {\csplits{\Theta}{\Theta_\Gamma}{\Theta_x}}
      {\issubst{\gamma}{\Theta}{\Gamma, \thyp{x}{A}{q}}}
    \\
    \boxed{\mc{L}: \ms{Label} \to \ms{Expr} \to \ms{Block} \text{ s.t. for fresh } x, [a/x]\mc{L}(\lbl{\ell}, x) = \mc{L}(\lbl{\ell}, a)}
    \\
    \prftree[r]{\rle{lb-subst-nil}}
    {\lbsubst{\mc{L}}{\cdot}{\ms{K}}}
      \qquad
    \prftree[r]{\rle{lb-subst-cons}}
      {\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
      {\haslb{\Gamma, \thyp{x}{A}{}}{\mc{L}(\lbl{\ell}, x)}{\ms{K}}}
      {\lbsubst{\mc{L}}{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K}}}
    % \\
    % \prftree[r]{\rle{lb-strict}}
    %   {\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
    %   {\ms{labels}(\ms{L}) = \ms{labels}(\ms{K})}
    %   {\strictlbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
  \end{gather*}
  \caption{Typing rules for \isotopessa substitutions. We assume that for all but finitely many \(x\), \(\lbl{\ell}\), \(\gamma(x) = x\) and \(\mc{L}(\lbl{\ell}, x) = \lbrb{\ell}{x}\).}
  \Description{Typing rules for isotope-SSA substitutions}
  \label{fig:ssa-subst-typing}
\end{figure}

\TODO{some text about substitution}

@@

\TODO{some text about substitution operations}

@@

\TODO{some text about substitution weakening}

\begin{theorem}[Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\)
    \item If \(\haslb{\Gamma}{t}{\ms{L}}\), then there exists \(\ms{K}\) such that \(\lbsubst{\substlbs{\gamma}{L}}{\ms{K}}{\ms{L}}\) and \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\)
  \end{itemize}
\end{theorem}

\TODO{some text about substitution}

@@

\TODO{some text about label substitution}

@@

\TODO{some text about terms with holes}

@@

\TODO{some text about term with hole typing rules}

@@

\TODO{some text about block with hole typing rules}

@@

\TODO{some text about rewrite typing rules}

\begin{theorem}[Rewriting]
  Given \(\isrw{\mc{H}}{H}{I}\), we have
  \begin{itemize}
    \item If \(\mhasty{H}{\Gamma}{p}{a}{A}{q}\), then \(\mhasty{I}{\Gamma}{p}{[\mc{H}]a}{A}{q}\)
    \item If \(\mhaslb{H}{\Gamma}{t}{\ms{L}}\), then \(\mhaslb{I}{\Gamma}{[\mc{H}]t}{\ms{L}}\)
  \end{itemize}

  Furthermore, we have that 
  \begin{itemize}
    \item \(\mhasty{\cdot}{\Gamma}{p}{a}{A}{q} \iff \hasty{\Gamma}{p}{a}{A}{q}\)
    \item \(\mhaslb{\cdot}{\Gamma}{t}{\ms{L}} \iff \haslb{\Gamma}{t}{\ms{L}}\)
  \end{itemize}
\end{theorem}
'
\TODO{some text about rewriting}

@@

\TODO{premonoidal/effectful categories}

@@

\TODO{relevant/affine/intuitionistic subcategories}

@@

\TODO{some text about type/context semantics}

@@

\TODO{some text about structural semantics}

@@

\TODO{some text about term semantics}

@@

\TODO{some text about block semantics}

\subsection{Semantic Metatheory}

\TODO{some text about semantic metatheory}

@@

\TODO{some text about coherence, implicit use above}

@@

\TODO{some text about upgrade}

\begin{lemma}[Semantic Weakening] 
  If \(\cwk{\Gamma}{\Delta}\), then 
  \begin{itemize}
    \item For all \(\hasty{\Delta}{p}{a}{A}{q}\),
    \[
      \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
      ; \dnt{\hasty{\Delta}{p}{a}{A}{q}}
      = \dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
    \]
    \item For all \(\haslb{\Delta}{t}{\ms{L}}\),
    \[
      \upg{\dnt{\cwk{\Gamma}{\Delta}}}{p}
      ; \dnt{\haslb{\Delta}{t}{\ms{L}}}
      = \dnt{\haslb{\Gamma}{t}{\ms{L}}}  
    \]
  \end{itemize}

  Similarly, if \(\lwk{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma}{t}{\ms{L}}\), then
  \[
    \dnt{\haslb{\Gamma}{t}{\ms{L}}}
    ; \dnt{\lwk{\ms{L}}{\ms{K}}}
    = \dnt{\haslb{\Gamma}{t}{\ms{K}}}
  \]
\end{lemma}

\TODO{some text about semantic weakening}

@@

\TODO{some text about substitution semantics}

\begin{lemma}[Substitution Coherence] \
  \begin{itemize}
    \item Given any two derivations \(D_1: \issubst{\gamma}{\Theta}{\Gamma}\), \(D_2: \issubst{\gamma'}{\Theta}{\Gamma}\) such that \(\forall x \in \Theta, \gamma(x) = \gamma'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
    \item Given any two derivations \(D_1: \lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), \(D_2: \lbsubst{\mc{L}'}{\ms{L}}{\ms{K}}\) such that \(\forall \lbl{\ell} \in \ms{L}, \mc{L}(x) = \mc{L}'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
    % \item Given any two derivations \(D_1: \strictlbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), \(D_2: \strictlbsubst{\mc{L}'}{\ms{L}}{\ms{K}}\) such that \(\forall \lbl{\ell} \in \ms{L}, \mc{L}(x) = \mc{L}'(x)\), we have \(\dnt{D_1} = \dnt{D_2}\)
  \end{itemize}
\end{lemma}

@@

\begin{theorem}[Semantic Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item For all \(\hasty{\Gamma}{p}{a}{A}{q}\), 
    \[
      \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{p}
      ;\dnt{\hasty{\Gamma}{p}{a}{A}{q}} 
      = \dnt{\hasty{\Theta}{p}{[\gamma]a}{A}{q}}
    \]
    \item For all \(\haslb{\Gamma}{t}{\ms{L}}\), \(\lbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}\), \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\),
    \[
      \upg{\dnt{\issubst{\gamma}{\Theta}{\Gamma}}}{}
      ; \dnt{\haslb{\Gamma}{t}{\ms{L}}}
      = \dnt{\haslb{\Theta}{[\gamma]t}{\ms{K}}} 
      ; \dnt{\lbsubst{\substlbs{\gamma}{\ms{K}}}{\ms{K}}{\ms{L}}}
    \]
  \end{itemize}
  Similarly, given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma}{t}{\ms{L}}\), we have
  \[
    \dnt{\haslb{\Gamma}{t}{\ms{L}}};\dnt{\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}}
    = \dnt{\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}}
  \]
\end{theorem}

\TODO{some text about semantic substitution}

@@

\TODO{some text about semantics of holes}

@@

\TODO{fill}

\TODO{some text about semantics of terms with holes}

@@

\TODO{fill}

\TODO{some text about semantics of blocks with holes}

@@

\TODO{fill}

\TODO{some text about semantics of rewrites}
